Forage site distributions quantify collective foraging in honey bee colonies

### 2022-02-08 analysis/landuse_effect_prop_scouts.Rmd
made changes recomended by Elli

@@

loading_plot <- ggplot(
  data = argi.rural.loadings,
  aes(
    x = Landuse, y = Correlation
  )
) +
  geom_boxplot(aes(fill = Landuse), lwd=0.1) +
  geom_hline(
    yintercept = 0, color = "black"
  ) +
  geom_text(
    data = rural_loading_text,
    aes(x = landuse_type, y = result, label = landuse_type),
    hjust = 0.5,
    vjust = 0.5,
    nudge_y = c(-0., -0.), nudge_x = c(0.5, 0.5)
  ) +
  annotate(
    "rect",
    ymin = -cutoff,
    ymax = cutoff,
    xmin = 0, xmax = Inf,
    alpha = 0.2
  ) +
  coord_flip() +
  theme_classic() +
  theme(
    axis.line.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "None",
    text = element_text()
  ) +
  scale_fill_manual(
    values = c(
      "Arable" = "#CCBB44", "Built Up Area" = "#EE6677",
      "Non Agricultural Unmanaged Green Space" = "#228833",
      "Water" = "#4477AA", Other = "#BBBBBB"
    )
  )

@@

loading_plot <- ggplot(
  data = urban.loadings,
  aes(
    x = Landuse, y = Correlation
  ),
  fill = Landuse
) +
  geom_boxplot(aes(fill = Landuse), lwd=0.1) +
  geom_hline(
    yintercept = 0, color = "black"
  ) +
  geom_text(
    data = urban_loading_text,
    aes(x = landuse_type, y = result, label = landuse_type),
    hjust = 0.5,
    vjust = 0.5,
    nudge_y = c(-0., -0.), nudge_x = c(0.5, 0.5)
  ) +
  annotate(
    "rect",
    ymin = -cutoff,
    ymax = cutoff,
    xmin = 0, xmax = Inf,
    alpha = 0.2
  ) +
  coord_flip() +
  theme_classic() +
  theme(
    axis.line.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "None",
    text = element_text()
  ) +
  scale_fill_manual(
    values = c(
      "Continuous Central" = "#BBBBBB", "Dense Residential" = "#BBBBBB",
      "Sparse Residential" = "#EE6677", "Railway" = "#CCBB44",
      "Parks Allotments Cemeteries" = "#228833",
      "Water" = "#BBBBBB", Other = "#BBBBBB"
    )
  )

>>>>

# Proportion of landuse in Urban and Agri-rural

```{r land-use-coverage, message = FALSE, warning = FALSE, error = TRUE}
rural_total_area <- rural_lu[, 3:ncol(rural_lu)] %>%
  colSums()
rural_total_area <- tibble(
  landuse_type = names(rural_total_area), area = rural_total_area
)
rural_total_area$area_pct <- (
  rural_total_area$area / sum(rural_total_area$area)
) * 100

rural_total_area <- rural_total_area %>%
  select(landuse_type, area_pct) %>%
  rename(
    Landuse = landuse_type, "% coverage" = area_pct
  ) %>%
  mutate(
    Landuse = tools::toTitleCase(
      gsub("\\.", " ", rural_total_area$landuse_type)
    ),
    Landuse = ifelse(Landuse == "OSR", "Oilseed Rape", Landuse)
  ) %>%
  rename(
    "Land-use" = Landuse
  )

rural_total_area

urban_total_area <- urban_lu[, 3:ncol(urban_lu)] %>%
  colSums()
urban_total_area <- tibble(
  landuse_type = names(urban_total_area), area = urban_total_area
)
urban_total_area$area_pct <- (
  urban_total_area$area / sum(urban_total_area$area)
) * 100

urban_total_area <- urban_total_area %>%
  select(landuse_type, area_pct) %>%
  rename(
    Landuse = landuse_type, "% coverage" = area_pct
  ) %>%
  mutate(
    Landuse = tools::toTitleCase(
      gsub("\\.", " ", urban_total_area$landuse_type)
    )
  ) %>%
  rename(
    "Land-use" = Landuse
  )

urban_total_area

save(urban_total_area, rural_total_area, file = "results/area_coverage.RData")
```

@@

loading_plot <- ggplot(
  data = argi.rural.loadings,
  aes(
    x = Landuse, y = Correlation
  )
) +
  geom_boxplot(aes(fill = Landuse), lwd = 0.1) +
  geom_hline(
    yintercept = 0, color = "black"
  ) +
  geom_text(
    data = rural_loading_text,
    aes(x = landuse_type, y = result, label = landuse_type),
    hjust = 0.5,
    vjust = 0.5,
    nudge_y = c(-0., -0.), nudge_x = c(0.5, 0.5), size = rel(textsize)
  ) +
  annotate(
    "rect",
    ymin = -cutoff,
    ymax = cutoff,
    xmin = 0, xmax = Inf,
    alpha = 0.2
  ) +
  coord_flip() +
  theme_classic() +
  theme(
    axis.line.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "None",
    text = element_text()
  ) +
  scale_fill_manual(
    values = c(
      "Arable" = "#CCBB44", "Built Up Area" = "#EE6677",
      "Non Agricultural Unmanaged Green Space" = "#228833",
      "Water" = "#4477AA", Other = "#BBBBBB"
    )
  )

@@

loading_plot <- ggplot(
  data = urban.loadings,
  aes(
    x = Landuse, y = Correlation
  ),
  fill = Landuse
) +
  geom_boxplot(aes(fill = Landuse), lwd = 0.1) +
  geom_hline(
    yintercept = 0, color = "black"
  ) +
  geom_text(
    data = urban_loading_text,
    aes(x = landuse_type, y = result, label = landuse_type),
    hjust = 0.5,
    vjust = 0.5,
    nudge_y = c(-0., -0.), nudge_x = c(0.5, 0.5), size = rel(textsize)
  ) +
  annotate(
    "rect",
    ymin = -cutoff,
    ymax = cutoff,
    xmin = 0, xmax = Inf,
    alpha = 0.2
  ) +
  coord_flip() +
  theme_classic() +
  theme(
    axis.line.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "None",
    text = element_text()
  ) +
  scale_fill_manual(
    values = c(
      "Continuous Central" = "#BBBBBB", "Dense Residential" = "#BBBBBB",
      "Sparse Residential" = "#EE6677", "Railway" = "#CCBB44",
      "Parks Allotments Cemeteries" = "#228833",
      "Water" = "#BBBBBB", Other = "#BBBBBB"
    )
  )

### 2022-02-08 analysis/manuscript/paper.Rmd
fixed box not rendering
made changes recomended by Elli

Honeybee foraging is a collective behaviour that is directed through the extraordinary waggle dance communication system, whereby colony foraging effort is allocated through a series of feedback loops [@Frisch1993; @Seeley1995]. Despite having been described as the most astounding example of non-primate communication [@Frisch1993], recent evidence suggests that the benefits of using dance information vary across environments and are sometimes even negligible [@Dornhaus2006; @Price2019]. Yet, because we cannot easily quantify collective foraging, pinpointing which environments promote collective foraging is challenging. Here, we present a mathematical model which allows inferring the proportion of 'scout' and 'recruit' dances on a hive's dance floor and quantify collective foraging. By applying this methodology to waggle dance data from twenty hives, we show that dance information may be used less in environments where resources are sparse, potentially because colonies are required to send out more scouts to find food in these circumstances. Our methodology provides a tool to quantify waggle dance use in undisturbed hives at a much larger scale than hitherto has been possible and to identify those ecological conditions that may have driven the evolution of this remarkable collective behaviour.

@@

Within honeybee colonies, a series of simple rules that determine when and how much bees perform the celebrated waggle dance (Box 1) mean that choices between feeding sites occur at the level of the group rather than the individual. For example, because the number of dance circuits performed by a bee on returning from a food source reflects the net energetic benefits of the trip, more of the colony's workforce will be recruited to the sweeter of two equidistant sources [@Seeley1991], or the closer of two equally sweet sources [@Hasenjager2022], without requiring any individual bee to compare options. This extraordinary system is a key example of how social insect colony behaviour can take on a form extending beyond that of the individual units [@Sumpter2006].

Despite such impressive complexity, in some situations, colonies that are prevented from communicating via dances achieve equal or even greater foraging success than their wild-type counterparts [@Sherman2002; @Dornhaus2006; @Gruter2013; @Price2019]. This indicates that bees frequently ignore the spatial information in dances.

The prevalent explanation for this apparent redundancy portrays dance-based collective foraging as an evolutionary anachronism that was useful in ancient tropical environments but offers little benefit in modern temperate environments [@Dornhaus2004; @Donaldson-Matasci2012; @Price2015]. However, fossil evidence has questioned the long-held assumption that the genus Apis evolved in tropical climates [@Kotthoff2013]. Moreover, given that dancing comes at a significant time cost [@Seeley1988], this raises the issue of why all temperate *Apis mellifera* colonies continue to dance. The dual questions of what foraging selection pressures drove the evolution of this uniquely complex collective behaviour, and why it continues to exist, are thus yet to be satisfactorily answered. Empirical studies have shown that honeybee colonies use dance information flexibly in response to their environment [@Sherman2002; @Dornhaus2004; @Dornhaus2006], suggesting that bees cease to rely on dance information if it provides little benefit [@Price2019]. Thus, identifying those environments whereby modern honeybee colonies forage collectively rather than through individual search is a key step towards understanding dance evolution [@Price2015].

---
**Box 1**

@@

Honeybees provide an ideal study system in which to evaluate collective foraging thanks to the famous “waggle dance”, which conveys information about location of a resource as a vector of distance (duration of the waggle run) and direction (angle of the dance) [@Frisch1993]. Through the observation and decoding of the waggle dance, a colony's dance floor provides a unique opportunity to eavesdrop on the communication and decision making leading to collective foraging decisions. Resource “quality” is conveyed through the number of waggle runs performed [@Frisch1993; @Boch1956; @Esch1961; @Seeley1994; @Seeley2000]). Although bees that follow dances do not specifically interpret this information on an individual level [@Hasenjager2021], the resulting over-representation of high-quality sites on the dance-floor means that they are more likely to encounter dances that advertise better forage [@Seeley1995], and provides the colony with a mechanism to select the most profitable resources in their environment [@Seeley1994A].

---

Overall, evidence that colony-level collective foraging offers any benefit over individual search has proved surprisingly hard to come by, mainly because this requires to distinguish scouts from recruits. Scouts are foragers that have located a resource through individual search behaviour, whilst recruits are foragers that have found a forage site by using the information provided in the dance of another bee [@Seeley1995]. Scouting and recruiting are not fixed behavioural categories, because individual bees can engage in both over the course of their foraging lifetime, and foragers can dance on return from any successful trip irrespective of whether they were recruited to the forage site or found it individually [@Beekman2007]. So far, scouts and recruits are distinguished by means of the foragers' individual search history, which requires following individual bees and is time and labour intensive [@Seeley1995].

@@

To reveal what the distribution of waggle run durations reported on the dance floor look like, and to demonstrate that dances from recruits and scouts lead to different distributions, we simulated honeybees foraging in a landscape where resource patches were randomly placed in the environment. Foragers could locate these under two different strategies: either acting as a scout and locating resources themselves, or following a recruit strategy and locating resources by following a random dance from the dance floor [@Seeley1992] (Fig. 1). As it is known in the simulation which individuals forage under what strategy, we can compare the distributions of foraging distances reported on the dance floor. Figs 1c and 1d show that the shapes of the resource distance distributions for the two types of foraging trips are very different. The distance distribution of the scouts is akin to that of an exponential distribution (Fig. 1 C), which is the nearest neighbour distance distribution for foragers operating in a one-dimensional environment (see methods). The distribution of the distances reported by recruits (Fig. 1 D) is a Rayleigh distribution which is the nearest-neighbour distribution of a two-dimensional environment [@Pyke1978] (see Methods).

@@

For 17 out of 20 study hives, a model of collective foraging provided a better explanation of the data than one of individual foraging (Fig. 3a). In each of these cases, the collective model had a good fit (using a Kolmogorov-Smirnov statistic, see methods) to the empirical waggle run durations, and for 9 of these sites, the scout model was significantly different to the data (Fig. 3b). The sites shown in Figs 3c-d are representative examples showing the model fits, showing the overall quality of the model fit to the data.

@@

Our results show that, whilst colony-level foraging mostly use a mixture of scouts and recruits, in some circumstances, colony foraging is better described by a model of individual foraging. This suggests that in some environments, honey bees actively ignore the waggle dance and instead prioritise the use of private information. Previous empirical studies have shown that individual foragers flexibly alter their use of waggle dance information in response to unrewarding resources [@Gruter2011; @Price2019]. Evaluating how these individual decisions influence the collective, however, has historically been a challenge as it is effectively impossible to track an individuals foraging behaviour over a landscape. As our results show that individual foraging accurately describes colony foraging in three different sites, these findings support the idea that individual decision making can dominate colony foraging and demonstrates further evidence for flexible waggle dance use by honey bee colonies.

Our finding that foraging in some landscapes is more accurately described by ignoring the dance and focusing on private information supports the hypothesis that the dance is used flexibly in response to environmental factors. As the collective model quantifies waggle dance use, the fit of our model provides a new tool to evaluate how environmental factors influence waggle dance use, and thus how a colony values waggle dance information under different environmental circumstances.

We compared the estimates of the proportion of scout and recruits for colonies in different environments. Each site used in this study is situated in a landscape comprised of different proportions of land-use types. To evaluate how these different land-use types influenced waggle dance use, we first classified the different land-use types of the area surrounding each site to obtain a standardised land-use profile for the urban and agri-rural environments separately (see methods). As many of the resulting land-use types covaried with each other, we performed a Partial Least Squares (PLS) analysis [@Carrascal2009] (see methods) to determine the principal components that represent combinations of land-use types which explained the most variation in the proportion of scout dances.

In the agri-rural environment one combination of land use types explained ~52% of the variation in the proportion of scouts (beta regression: $R^2 = 0.52$, $\phi$ = 3.7, p < 0.05, Fig. 4. A). This principal component correlates positively with arable land and negatively with non-agricultural unmanaged green space, built-up areas and water (Fig. 4. B). From the land-use maps coloured by each landscape dominating PC1 (Fig 4) we can see that arable land and built-up areas show the most coverage. As arable land increases whilst unmanged green space and built-up areas decrease, the proportion of individuals prioritising private information increases.

@@

A similar effect is also observed in the urban environment, where a single PC explained ~59% of the variance in the proportion of scouts (beta regression: $R^2$ = 0.59, $\phi$ = 6.2, p < 0.05, Fig. 5. A). This PC correlates positively with parks, allotments and cemeteries, whilst correlating negatively with sparse residential lands and railways (Fig. 5. B). Visualising how waggle dance use is influenced by changes in these land-use types shows that when the landscape is dominated by sparse residential land, as in site HHS, waggle dance use dominates the colonies foraging strategy. However, this reduces as sparse residential land is replaced by parks, allotments, and cemeteries (Fig. 5).

@@

In agri-rural environments, land-use types which typically associate with resource scarcity correlate positively with a higher proportion of scouts, whilst those which correlate with a higher resource abundance correlate with a lower proportion of scouts. Arable land is typically nutritionally poor for honey bees [@Samuelson2021]. On the contrary, non-agricultural unmanaged green space is typically nutrient rich, as are built up areas, which often have gardens with ample flowers available for pollinators [@Baldock2019]. These findings indicate that in agri-rural environments, when resources are more abundant and thus easier to find, the collective foraging is used more, presumably to identify the best resources in the environment. Conversely, where resources are sparse and more difficult to find, a decreased use of the waggle dance suggests the colony focuses more on individual exploration rather than in trying to locate on the best resources. A similar result is observed for urban environments, where the dependence appears to be driven by the presence of flower rich gardens in residential areas. However, an increase in the amount of parks, allotments and cemeteries appears to decrease the use of the dance.

( **NOTE:** I feel like we need more of an explanation here. For the agri-rural environment we can assemble the results into a coherent story which has some logical basis and suggests more data would investigate this further. The results for the urban sites are contradictory. Parks, allotments and cemeteries increases the proportion of scouts so following our logic these areas should be barron, but the evidence suggests they are not and along side gardens should be full of flora. This indicates there is something else in play here, such as the spatial arrangement of the locations? Or that we just don't have enough data to comment on what processes are going on in urban sites. However, we do get a 'significant' result in the PLS. Maybe we need more fine grained data to evaluate how season influences, it could be that these results are muddled by an interaction with season or temperature or something. Also, railway is a random thing to pop up as significant. I am not sure how best to describe these results without it looking like blatant data mining...)

>>>>

Honeybee foraging is a collective behaviour that is directed through the extraordinary waggle dance communication system, whereby colony foraging effort is allocated through a series of feedback loops [@Frisch1993; @Seeley1995]. Despite having been described as the most astounding example of non-primate communication [@Frisch1993], recent evidence suggests that the benefits of using dance information vary across environments and are sometimes even negligible [@Dornhaus2006; @Price2019]. Yet, because we cannot easily quantify collective foraging, pinpointing which environments promote collective foraging is challenging. Here, we present a mathematical model which quantifies collective foraging by inferring the proportion of 'scout' and 'recruit' dances on a hive's dance floor. By applying this methodology to waggle dance data from twenty hives, we show that dance information may be used less in environments where resources are sparse, potentially because colonies are required to send out more scouts to find food in these circumstances. Our methodology provides a tool to quantify waggle dance use in undisturbed hives at a much larger scale than hitherto has been possible and to identify those ecological conditions that may have driven the evolution of this remarkable collective behaviour.

@@

Within honeybee colonies, a series of simple rules that determine when and how much bees perform the celebrated waggle dance (Figure of dance waggle dance) mean that choices between feeding sites occur at the level of the group rather than the individual. For example, because the number of dance circuits performed by a bee on returning from a food source reflects the net energetic benefits of the trip, more of the colony's workforce will be recruited to the sweeter of two equidistant sources [@Seeley1991], or the closer of two equally sweet sources [@Hasenjager2022], without requiring any individual bee to compare options. This extraordinary system is a key example of how social insect colony behaviour can take on a form extending beyond that of the individual units [@Sumpter2006].

Despite such impressive complexity, in some situations, colonies that are prevented from communicating via dances achieve equal or even greater foraging success than their wild-type counterparts [@Sherman2002; @Dornhaus2006; @Gruter2013; @Price2019]. This indicates that using the dance does not always provide a foraging advantage and suggests bees may ignore spatial information in dances in favour of independent foraging. The prevalent explanation for this apparent redundancy portrays dance-based collective foraging as an evolutionary anachronism that was useful in ancient tropical environments but offers little benefit in modern temperate environments [@Dornhaus2004; @Donaldson-Matasci2012; @Price2015].

Given that dancing comes at a significant time cost [@Seeley1988], this raises the issue of why all temperate *Apis mellifera* colonies continue to dance. The dual questions of what foraging selection pressures drove the evolution of this uniquely complex collective behaviour, and why it continues to exist, are thus yet to be satisfactorily answered. Empirical studies have shown that honeybee colonies use dance information flexibly in response to their environment [@Sherman2002; @Dornhaus2004; @Dornhaus2006], suggesting that bees cease to rely on dance information if it provides little benefit [@Price2019]. Identifying those environments whereby modern honeybee colonies forage collectively rather than through individual search is a key step towards understanding dance evolution [@Price2015].

**Box 1**

@@

Honeybees provide an ideal study system in which to evaluate collective foraging thanks to the famous “waggle dance”, which conveys information about location of a resource as a vector of distance (duration of the waggle run) and direction (angle of the dance relative to the sun) [@Frisch1993].Through the observation and decoding of the waggle dance, a colony's dance floor provides a unique opportunity to eavesdrop on the communication and decision making leading to collective foraging decisions. Resource “quality” is provided through the number of waggle runs performed [@Frisch1993; @Boch1956; @Esch1961; @Seeley1994; @Seeley2000]). Although bees that follow dances do not specifically interpret this information on an individual level [@Seeley1991], the resulting over-representation of high-quality sites on the dance-floor means that they are more likely to encounter dances that advertise better forage [@Seeley1995], and provides the colony with a mechanism to select the most profitable resources in their environment [@Seeley1994A].

**End Box 1**

Overall, evidence that colony-level collective foraging offers any benefit over individual search has proved surprisingly hard to come by. Whilst it is entirely possible the dance no longer provides a benefit in the modern world, a key barrier to assessing waggle dance use is that this requires distinguishing between foraging trips made by
scouting and recruited individuals. Scout foraging trips occur when a food source is located through individual search behaviour, whilst recruit foraging trips occur when a forage site is found by using the information provided in the dance of another bee [@Seeley1995]. Scouting and recruiting are not fixed behavioural categories, because individual bees can engage in both over the course of their foraging lifetime, and foragers can dance on return from any successful trip irrespective of whether they were recruited to the forage site or found it individually [@Beekman2007]. So far, scouts and recruits are distinguished by means of the foragers' individual search history, which requires following individual bees and is time and labour intensive [@Seeley1995].

@@

To reveal how the patterns of waggle dance encoded foraging distances might differ under a scout and recruit foraging strategy, we simulated honeybees foraging in a landscape where resource patches were randomly placed in the environment. Foragers could locate these under two different strategies: either acting as a scout and locating resources themselves, or following a recruit strategy and locating resources by following a random dance from the dance floor [@Seeley1992] (Fig. 1). As it is known in the simulation which individuals forage under what strategy, we can compare the distributions of foraging distances reported on the dance floor. Figs 1c and 1d show that the shapes of the resource distance distributions for the two types of foraging trips are very different. The distance distribution of the scouts is akin to that of an exponential distribution (Fig. 1 C), which is the nearest neighbour distance distribution for foragers operating in a one-dimensional environment (see methods). The distribution of the distances reported by recruits (Fig. 1 D) is a Rayleigh distribution which is the nearest-neighbour distribution of a two-dimensional environment [@Pyke1978] (see Methods).

@@

For 17 out of 20 study hives, a model of collective foraging provided a better explanation of the data than one of individual foraging (Fig. 3a). In all but one of these cases, the collective model had a good fit (using a Kolmogorov-Smirnov statistic, see methods) to the empirical waggle run durations, whereas for 9 of these sites, the distribution predicted by the scout model was significantly different to the observed data (Fig. 3b). The sites shown in Figs 3c-d are representative examples showing the model fits, showing the overall quality of the model fit to the data.

@@

Our results show that, whilst colony-level foraging is mostly comprised of a mixture of scout and recruit foraging trips, in some circumstances, colony foraging can be better described by individual foraging alone. This suggests that in some environments honey bees engage in scouting to find new food sites rather than follow dances. Previous empirical studies have shown that individual foragers flexibly alter their use of waggle dance information in response to unrewarding resources [@Gruter2011; @Price2019]. Evaluating how these individual decisions influence the collective, however, has historically been a challenge as it is effectively impossible to track an individuals foraging behaviour over a landscape. As our results show that individual foraging accurately describes colony foraging in three different sites, these findings support the idea that individual decision making can dominate colony foraging and demonstrates further evidence for flexible waggle dance use by honey bee colonies.

Our finding that foraging in some landscapes is more accurately described by scouting supports the hypothesis that the dance is used flexibly in response to environmental factors. As the collective model quantifies the contribution of recruitment to colony foraging, the fit of our model provides a new tool to evaluate how environmental factors influence waggle dance use, and thus how a colony uses waggle dance information under different environmental circumstances.

We estimated the proportion of scout and recruit foraging trips for each individual colony and investigated whether these estimates might correlate with land-use. Each site used in this study is situated in a landscape comprised of different proportions of land-use types. To evaluate how these different land-use types influenced waggle dance use, we first classified the different land-use types of the area surrounding each site to obtain a standardised land-use profile for the urban and agri-rural environments separately as many land-use types present in urban areas do not occur in agri-rural environments and vice versa (see methods). As many of the resulting land-use types covaried with each other, we performed a Partial Least Squares (PLS) analysis [@Carrascal2009] (see methods) to determine the principal components that represent combinations of land-use types which explained the most variation in the proportion of scout dances within agri-rural and urban environments. As we have a small sample size, this analysis was jackknifed to evaluate the robustness of our results to influential points (see methods, supplementary material section _._)

In the agri-rural environment one combination of land use types explained ~52% of the variation in the proportion of scouts (beta regression: $R^2 = 0.52$, $\phi$ = 3.7, p < 0.05, Fig. 4. A). This principal component correlates positively with arable land and negatively with non-agricultural unmanaged green space, built-up areas and water (Fig. 4. B). These land-use types maintain a significant correlation with PC1 over the jackknifed PLS (Fig. 4. B) indicating the results are robust. As arable land increases whilst unmanaged green space and built-up areas decrease, the proportion of individuals prioritising private information increases. Arable land dominates the agri-rural environment accounting for over 28% of land-coverage, whilst built-up areas represent around 17% of lands (Table 1). Non-agricultural unmanaged green space represent smaller amounts of land, accounting for less than 3% of the total land combined (Table 1).

```{r, agri-rural-area-table}
set_flextable_defaults(
  big.mark = " ",
  font.size = 10, theme_fun = theme_vanilla,
  padding.bottom = 6,
  padding.top = 6,
  padding.left = 6,
  padding.right = 6
)

load("../results/area_coverage.RData")

rural_total_area %>%
  arrange(desc(`% coverage`)) %>%
  mutate(`% coverage` = signif(`% coverage`, 3)) %>%
  flextable() %>%
  set_caption(
    "Percentage area covered for each land-use type in the agri-rural environment"
  ) %>%
  set_table_properties(layout = "autofit")
```

@@

A similar effect is also observed in the urban environment, where a single PC explained ~59% of the variance in the proportion of scouts (beta regression: $R^2$ = 0.59, $\phi$ = 6.2, p < 0.05, Fig. 5. A). This PC correlates positively with parks, allotments and cemeteries, whilst correlating negatively with sparse residential lands and railways (Fig. 5. B). However, the jackknifed PLS identifies CAD as a highly influential site, the removal of which causes sparse residential land to have a positive correlation with PC1 (Supplementary Material _._). Visualising how waggle dance use is influenced by changes in these land-use types shows that when the landscape is dominated by sparse residential land, as in site HHS, waggle dance use dominates the colonies foraging strategy. However, this reduces as sparse residential land is replaced by parks, allotments, and cemeteries (Fig. 5). Sparse residential land dominates the urban environment accounting for over 34% of land-coverage, whilst parks allotments and cemeteries and railway corresponds to ~8% and ~1 respectively (Table 2).

```{r, urban-area-table}
urban_total_area %>%
  arrange(desc(`% coverage`)) %>%
  mutate(`% coverage` = signif(`% coverage`, 3)) %>%
  flextable() %>%
  set_caption(
    "Percentage area covered for each land-use type in the agri-rural environment"
  ) %>%
  set_table_properties(layout = "autofit")
```

@@

In agri-rural environments, land-use types which typically associate with resource scarcity correlate positively with a higher proportion of scouts, whilst those which correlate with a higher resource abundance correlate with a lower proportion of scouts. Arable land is typically nutritionally poor for honey bees [@Samuelson2021]. On the contrary, built up areas often have gardens with ample flowers available for pollinators [@Baldock2019]. Although non-agricultural unmanaged green space and water show as significantly correlating with reduced use of the dance, as they cover such a small amount of area, combined with a small sample size of only 10 sites, we cannot ignore the possibility that their significance is a random effect rather than evidence of a biological mechanism. Overall, these findings indicate that in agri-rural environments, when resources are more abundant and thus easier to find, the waggle dance is used more, presumably to identify the best resources of those on offer. Conversely, where resources are sparse and more difficult to find, a decreased use of the waggle dance suggests the colony focuses more on individual exploration rather than in trying to locate the best resources.

A similar result is observed in urban sites, where the presence of flower rich gardens in residential areas appears to promote increased use of the dance. Conversely, an increase in the amount of parks, allotments and cemeteries appears to decrease the use of the dance. Such land-use types could be expected to be quite resource rich for pollinators yet our findings suggests honey bee colonies decrease the use of the dance in such environments. It is possible that such environments vary seasonally in terms of their forage availability, allowing them to function both as floral deserts and oasis at different times of the year. As our waggle dance observations were collected over the course of an entire season, we may only be picking up the dominant signal over this time frame. However, the jackknife analysis shows our results are highly sensitive to influential points. Further data is therefore needed to investigate how the urban environment influences waggle dance use in honey bees.
